---
title: "Movies Dataset"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Exploratory Data Analysis Project
#### By Michael Bohn, Nate Wagner, Dominic Ventura


```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
#setwd("/Users/natewagner/Documents/Data_munging_eda_project")
```



```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(readr)
library(tidyverse)
library(lubridate)
library(jsonlite)
library(data.table)
library(tm)
library(wordcloud)
library(RColorBrewer)
library(knitr)
library(gridExtra)
library(ggforce)
library(SnowballC)
library(plotly)
#install.packages("gt")
#install.packages("tm")
#install.packages("wordcloud")
#install.packages("RColorBrewer")
```



```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
movies <- read_csv("movies_metadata.csv")
movies.orig <- read_csv("movies_metadata.csv")
credits <- read_csv("credits.csv")
keywords <- read_csv("keywords.csv")
links <- read_csv("links.csv")
ratings <- read_csv("ratings.csv")
cast_data <- read_csv("castData.csv")
cast_data1 <- read_csv("castDataCrew.csv")
production <- read_csv("Production.csv")
metascrub <- read_csv("MetaScrubbedMoney.csv")
metascrubdf <- as.data.frame(metascrub)
```


### Completeness

Most of our variables of interest were complete.
```{r echo=FALSE, message=FALSE, warning=FALSE}
# get average rating for other users
avg_ratings <- select(ratings, movieId, rating)  %>% group_by(movieId) %>% summarise_at(vars(rating), list(rating = mean))
names(avg_ratings) <- c("id", "avg_rating")

# join tables:
names(links) <- c("id", "imdbid", "tmdbid")
movies.merged <- merge(x = links, y = avg_ratings, by = "id")
movies.merged1 <- merge(x = movies, y = select(movies.merged, tmdbid, avg_rating), by.x = "id", by.y = "tmdbid")
movies <- movies.merged1


sum.NA.total <- movies %>%
  select(everything()) %>%
  summarise_all(funs(sum(is.na(.))))

PropOfNA.total <- (sum.NA.total/nrow(movies))*100
SumNA.prop <- PropOfNA.total %>% map_dbl(sum)
kable(SumNA.prop, format = "markdown")

```


### Exploratory Data Analysis

We have data on movies from 1874 to 2020. The number of movies per year clearly increases over time. 
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Year Summary
sum_year <- as.data.frame(as.array(summary(year(movies$release_date))))
names(sum_year) <- c("Statistic", "Value")
kable(sum_year, type = "Markdown")

# freq table for # movies per year:
movies_byYear <- movies %>% 
  group_by(year(release_date)) %>% 
  count()

# barplot of # movies per year:
ggplot(movies_byYear, aes(`year(release_date)`, n)) + 
  geom_bar(stat="identity", width = 0.5, fill = "tomato2") + 
  labs(title = "Number of Movies per Year", x = "Year", y = "Number of Movies") + 
  theme_classic()

```


### What affects movie ratings?


We have two populations of user's movie ratings. With the "avg_rating" in movies corresponding to ratings of MovieLens users, and "vote_average" corresponding to TMDb users.
There are a couple extreme outliers with average ratings of 0 and vote counts greater than 30. For simplicity and to get a better representation of our scatterplots we are going to exclude these two points. 
```{r echo=FALSE, message=FALSE, warning=FALSE}
# TMDb users
sum_vote <- as.data.frame(as.array(summary(movies$vote_average)))
names(sum_vote) <- c("Vote Average for TMDb Users", "Value")
#kable(sum_vote, type = "Markdown")

# IMDb users
sum_vote2 <- as.data.frame(as.array(summary(movies$avg_rating)))
names(sum_vote2) <- c("Vote Average for MovieLens Users", "Value")
#kable(sum_vote2, type = "Markdown")

# vote count TMDb users
sum_vote3 <- as.data.frame(as.array(summary(movies$vote_count)))
names(sum_vote3) <- c("Vote Count", "Value")
kable(sum_vote3, type = "Markdown")


#movies %>% filter(vote_average < .50 & vote_count > 20) %>%
#  select(title, vote_average, vote_count)

movies <- movies %>%
  filter(vote_average > 0.50)
```


####Correlation Matrix
```{r echo=FALSE, message=FALSE, warning=FALSE}
# select numeric columns & add year column:
movies <- mutate(movies, year = year(movies$release_date))
metascrubdf <- mutate(metascrubdf, year = year(as.Date(metascrubdf$release_date)))

num.cols <- dplyr::select_if(movies, is.numeric)

# compute cor from all numeric columns:
cor.matrix <- round(cor(num.cols, use = "pairwise"), 2)

# melt the cor matrix:
cor.matrix.melt <- melt(cor.matrix)
```


Looking at the correlation matrix, there doesn't seem to be any real strong linear relationships between any of these variables and average movie ratings. But perhaps there are some non-linear relationships. 
```{r echo=FALSE, message=FALSE, warning=FALSE}
# plot cor matrix
ggplot(data = cor.matrix.melt, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() + theme(axis.text.x = element_text(angle=65, vjust=0.6))

kable(cor.matrix, format = "markdown")
```


Density plots for popularity, vote count, runtime and average vote.
```{r echo=FALSE, message=FALSE, warning=FALSE}
one <- ggplot(movies %>% filter(popularity < 20), aes(popularity)) + geom_density(fill="gray30", alpha=0.4, adjust=0.05) + theme_bw()

two <- ggplot(movies %>% filter(vote_count < 100 & vote_count > 0), aes(vote_count)) + geom_density(fill="gray30", alpha=0.4, adjust=0.4) + theme_bw()

three <- ggplot(movies %>% filter(runtime < 140 & runtime > 0), aes(runtime)) + geom_density(fill="gray30", alpha = 0.4, adjust = 0.3) + theme_bw()

four <- ggplot(movies, aes(vote_average)) + geom_density(fill="gray30", alpha=0.4, adjust = 0.35) + theme_bw()

grid.arrange(one, two, three, four, nrow = 2)
```



We are going to assume that to have an unbiased estimate of a movie's true average rating, there must be at least 30 votes, and we find 12,439 movies that meet this criteria. However, a potential source of bias with this approach is that it could be that movies with very low vote totals, are not very good movies to begin with and aren't popular. Thus, that could be why they have low vote counts and possible low average ratings.  
```{r message=FALSE, warning=FALSE, include=FALSE}
movies %>% 
  filter(vote_count >= 30) %>%
  count()
```


### Are low vote counts associated with lower average ratings?

It could be hard to see the actual trend with the extreme values of vote count.
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(movies %>% filter(vote_count > 0), aes(vote_count, vote_average)) +
  geom_point(alpha = 0.05) + 
  geom_smooth() + 
  labs(title = "Vote Count vs Vote Average", x = "Vote Count", y = "Vote Average") +
  theme_bw()
```


Even with removing the extreme values of vote count, it's still hard to see much of a releationship between vote count and vote average. 
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(movies %>% filter(vote_count > 0 & vote_count < 31), aes(vote_count, vote_average)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  labs(title = "Vote Count vs Vote Average", subtitle = "n < 31", x = "Vote Count", y = "Vote Average") +
  theme_bw()
```



```{r echo=FALSE, message=FALSE, warning=FALSE}
# Remove movies with vote count less than 30
movies <- movies %>% filter(vote_count >= 30)
```


### How has average vote changed over time?

It seems there is a slight negative relationship between average vote and when the movie was released. 
```{r echo=FALSE, message=FALSE, warning=FALSE}
p1 <- ggplot(movies, aes(year, vote_average)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  labs(title = "Release Date vs Vote Average", subtitle = "TMDb Users",x = "Year", y = "Vote Average") + 
  theme_bw()

p2 <- ggplot(movies, aes(year, avg_rating)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  labs(title = "Release Date vs Vote Average", subtitle = "MovieLens Users",x = "Year", y = "Vote Average") + 
  theme_bw()

grid.arrange(p1, p2, nrow = 1, ncol = 2)
```

### How does the runtime of a movie affect average votes?

```{r echo=FALSE, message=FALSE, warning=FALSE}
sum_runtime <- as.data.frame(as.array(summary(movies$runtime)))
names(sum_runtime) <- c("Runtime", "Value")
#kable(sum_runtime, type = "Markdown")

#movies %>%
#  filter(runtime == 0) %>%
#  count()
```

Hard to see association with the extreme values of runtime. However, even with the removal of outliers it's still hard to see much of an association between runtime and average movie rating.
```{r echo=FALSE, message=FALSE, warning=FALSE}
lower <- quantile(movies$runtime, 0.25, na.rm = TRUE) - 1.5 * IQR(movies$runtime, na.rm =T)
upper <- quantile(movies$runtime, 0.75, na.rm = TRUE) + 1.5 * IQR(movies$runtime, na.rm =T)

P1.1 <- ggplot(movies %>% filter(runtime > 0), aes(runtime, vote_average)) +
          geom_point(alpha = 0.1) +
          facet_zoom(x = runtime >= lower & runtime <= upper & runtime > 0) +
          geom_smooth() +
          theme_bw() +
          labs(title = "Runtime vs Vote Average", subtitle = "TMDb Users",x = "Runtime", y = "Vote Average")

P1.2 <- ggplot(movies %>% filter(runtime > 0), aes(runtime, avg_rating)) +
          geom_point(alpha = 0.1) +
          facet_zoom(x = runtime >= lower & runtime <= upper & runtime > 0) +
          geom_smooth() +
          theme_bw() +
          labs(title = "Runtime vs Vote Average", subtitle = "MovieLens Users",x = "Runtime", y = "Vote Average")

grid.arrange(P1.1, P1.2, nrow = 1, ncol = 2)
```



### Does average votes increase with popularity? 

```{r echo=FALSE, message=FALSE, warning=FALSE}
sum_popularity <- as.data.frame(as.array(summary(movies$popularity)))
names(sum_popularity) <- c("Popularity", "Value")
#kable(sum_popularity, type = "Markdown")
```

Even with the removal of extreme outliers, there seems to be no relationship between movie popularity and average movie rating.
```{r echo=FALSE, message=FALSE, warning=FALSE}
lower <- quantile(movies$popularity, 0.25, na.rm = TRUE) - 1.5 * IQR(movies$popularity, na.rm =T)
upper <- quantile(movies$popularity, 0.75, na.rm = TRUE) + 1.5 * IQR(movies$popularity, na.rm =T)

P1.3 <- ggplot(movies %>% filter(popularity > 0), aes(popularity, vote_average)) +
          geom_point(alpha = 0.1) +
          facet_zoom(x = popularity >= lower & popularity < upper & popularity > 0) +
          geom_smooth() +
          theme_bw() +
          labs(title = "Popularity vs Vote Average", subtitle = "TMDb Users",x = "Popularity", y = "Vote Average")

P2.3 <- ggplot(movies %>% filter(popularity > 0), aes(popularity, avg_rating)) +
          geom_point(alpha = 0.1) +
          facet_zoom(x = popularity >= lower & popularity < upper & popularity > 0) +
          geom_smooth() +
          theme_bw() +
          labs(title = "Popularity vs Vote Average", subtitle = "MovieLens Users",x = "Popularity", y = "Vote Average")


grid.arrange(P1.3, P2.3, nrow = 1, ncol = 2)
```


Another view of these relationships.
```{r echo=FALSE, message=FALSE, warning=FALSE}
tm <- plot_ly(movies %>% filter(popularity >= lower & popularity < upper & popularity > 0), x=~year, y=~vote_count, z=~vote_average,color=~popularity, mode= "markers",type = "scatter3d") %>% layout(title = "Vote Average vs Vote Count and Year", scene = list(
      xaxis = list(title = "Year"),
      yaxis = list(title = "Vote Count"),
      zaxis = list(title = "Vote Average")))
tm
```
It seems there could be some non-linear relationship between vote count and vote average here. 


### Word association in movies: 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Create the corpus of data
keywordsDS.corpus <- Corpus(VectorSource(keywords$keywords))
# Cleaning dat. This includes removing stopwords, numbers, whitespace, etc. and converting the corpus into a plain text document.
keywordsDS.Clean <- tm_map(keywordsDS.corpus, PlainTextDocument)
keywordsDS.Clean <- tm_map(keywordsDS.corpus, tolower)
keywordsDS.Clean <- tm_map(keywordsDS.Clean, removeNumbers)
keywordsDS.Clean <- tm_map(keywordsDS.Clean, removeWords, stopwords("english"))
keywordsDS.Clean <- tm_map(keywordsDS.Clean, removeWords, c("name"))
keywordsDS.Clean <- tm_map(keywordsDS.Clean, removeWords, c("film"))
keywordsDS.Clean <- tm_map(keywordsDS.Clean, removeWords, c("director"))
keywordsDS.Clean <- tm_map(keywordsDS.Clean, removePunctuation)
keywordsDS.Clean <- tm_map(keywordsDS.Clean, stripWhitespace)
keywordsDS.Clean <- tm_map(keywordsDS.Clean, stemDocument)
wordcloud(words = keywordsDS.Clean, min.freq = 1,
          max.words=100, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
```


The top three most common words are woman, relationship and independent.
```{r echo=FALSE, message=FALSE, warning=FALSE}

# Build a term-document matrix. Document matrix is a table containing the frequency of the words. Column names are words and row names are documents.
dtm <- TermDocumentMatrix(keywordsDS.Clean)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
names(d) <- c("Word", "Frequency")
kable(head(d, 10), type = "Markdown")


# Analyze the association between frequent terms (i.e., terms which correlate)
#findAssocs(dtm, terms = "love", corlimit = 0.3)
#findAssocs(dtm, terms = "music", corlimit = 0.3)
#findAssocs(dtm, terms = "war", corlimit = 0.3)
#findAssocs(dtm, terms = "world", corlimit = 0.3)


# barplot of Most Frequent words 
ggplot(d[1:20,], aes(reorder(Word, -Frequency), Frequency)) +
  geom_bar(stat="identity", width = 0.5, fill = "navy") +
  labs(title = "Most frequent words", y = "Word frequencies", x = "Words") + 
  theme(axis.text.x=element_text(angle=55, hjust=1))


```


### Breakdown of movies by cast gender:
```{r echo=FALSE, message=FALSE, warning=FALSE}
cast_data$gender[cast_data$gender == 0] <- "Unspecified"
cast_data$gender[cast_data$gender == 1] <- "Female"
cast_data$gender[cast_data$gender == 2] <- "Male"

ggplot(cast_data %>% filter(gender != "Unspecified"),
  aes(factor(gender), fill=factor(gender))) +
  geom_bar() +
  labs(title= "Movies by Gender", x = "Gender", y = "Count") +
  theme_bw()
```


### Top Billing by Actor
```{r echo=FALSE, message=FALSE, warning=FALSE}
order01 <- cast_data %>%
  filter(order == 0 | order == 1)

topBilling <- order01 %>%
  distinct() %>%
  count(name)
names(topBilling) <- c("Name", "Count")
topBilling <- topBilling[order(-topBilling$Count),]
#kable(head(topBilling,10), type = "Markdown")
  
topBilling10 <- head(topBilling[order(-topBilling$Count),], 10)
ggplot(data=topBilling10, aes(x = reorder(Name, -Count), y = Count)) +
  geom_col(aes(fill=Name)) + 
  theme(axis.text.x=element_text(angle=55, hjust=1)) +
  labs(x = "Actor", y = "Count")
```

### Top Billing by Gender
```{r echo=FALSE, message=FALSE, warning=FALSE}
billingGender <- order01 %>%
  filter(order == 0 | order == 1) %>%
  count(gender)

ggplot(billingGender %>% filter(gender != "Unspecified"), aes(x = reorder(gender, -n), y = n)) + geom_col(aes(fill=gender)) +
  labs(x = "Gender", y = "Count") +
  theme_bw()
```


### Production Companies: Revenue and Expenses
```{r echo=FALSE, message=FALSE, warning=FALSE}

comps <- aggregate(revenue ~ production_companies, data=production, FUN=sum) %>% 
  filter(revenue > 10000000000)

names(comps) <- c("Production Companies", "Revenue")
#kable(comps[order(-comps$Revenue),], type = "Markdown")

comps.budget <- aggregate(budget ~ production_companies, data=production, FUN=sum) %>%
  filter(budget > 2000000000)

new <- merge(x = comps, y = comps.budget, by.x = "Production Companies", by.y = "production_companies", all.x = TRUE)

names(new) <- c("Production Company", "Revenue", "Budget")

p <- plot_ly() %>%
  add_bars(
    x = c("Columbia Pictures", "New Line Cinema", "Paramount Pictures", "Twentieth Century Fox Film Corporation", "Universal Pictures", "Walt Disney Pictures", "Warner Bros."),
    y = c(10959150943,5433685000,12386353473,7132835801,12442067806,11940137358,3824898082),
    base = c(-10959150943,-5433685000,-12386353473,-7132835801,-12442067806,-11940137358,-3824898082),
    marker = list(
      color = 'red'
    ),
    name = 'expenses'
  ) %>%
  add_bars(
    x = c("Columbia Pictures", "New Line Cinema", "Paramount Pictures", "Twentieth Century Fox Film Corporation", "Universal Pictures", "Walt Disney Pictures", "Warner Bros."),
    y = c(27977585562,14456093103,37409828205,24203630979,41215102933,38536178369,12270573084),
    base = 0,
    marker = list(
      color = 'green'
    ),
    name = 'revenue'
  ) 

p
```



### How has revenue changed over time? 

Here we have two different views of the association. 
```{r echo=FALSE, message=FALSE, warning=FALSE}
PP1 <- ggplot(movies %>% filter(revenue > 100000000), aes(release_date, revenue)) +
        geom_point(alpha = 0.3) +
        theme_bw() +
        geom_smooth() +
        labs(title = "Revenue vs Release Date", x = "Release Date", y = "Revenue")
 
rev_by_year <- select(movies %>% filter(revenue > 0), year, revenue) %>% group_by(year) %>% summarise(Mean = mean(revenue, na.rm = TRUE))


PP2 <- ggplot(rev_by_year, aes(year, Mean)) + 
        geom_bar(stat="identity", width = 0.5, fill = "tomato2") + 
        labs(title = "Average Revenue per Year", x = "Release Date", y = "Average Revenue") + 
        theme_classic()

grid.arrange(PP1, PP2, nrow=1)
```


### Revenue vs Popularity
```{r echo=FALSE, message=FALSE, warning=FALSE}
pop.rev <- ggplot(movies %>% filter(revenue > 100000000), aes(popularity, revenue)) + 
            geom_point(alpha = 0.2) +
            geom_smooth() +
            labs(title = "Popularity vs Vote Average", x = "Popularity", y = "Revenue") +
            theme_bw()

ggplotly(pop.rev)
```


#### Average budget, revenue, and profit:
```{r echo=FALSE, message=FALSE, warning=FALSE}
rev <- sapply(metascrubdf[,c(4,5,6)],mean)
names(rev) <- c("Budget", "Revenue", "Profit")
kable(rev, type = "Markdown")
```

#### Correlations
```{r echo=FALSE, message=FALSE, warning=FALSE}
cormat <- cor(metascrubdf[,c(3:6,8)])
kable(cormat, type = "Markdown")
```


### Revenue vs Budget

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(metascrubdf, aes(budgetM, revenueM)) +
  geom_point(alpha = 0.2) +
  geom_smooth() +
  labs(title = "Revenue vs Budget", x = "Budget", y = "Revenue") +
  theme_bw()
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
movies0 <- movies %>% filter(budget > 0 & revenue > 0)

PP <- plot_ly(movies0 %>% filter(popularity >= lower & popularity < upper & popularity > 0), x=~budget, y=~year, z=~revenue,color=~popularity, mode= "markers",type = "scatter3d") %>% layout(title = "Revenue vs Year and Budget", scene = list(
      xaxis = list(title = "Budget"),
      yaxis = list(title = "Year"),
      zaxis = list(title = "Revenue")))
PP
```

#### Two main outliers:
```{r echo=FALSE, message=FALSE, warning=FALSE}
maxb <- which.max((metascrubdf$budgetM))
maxb <- metascrubdf[maxb,c(1,4,5)]
names(maxb) <- c("Title", "Budget", "Revenue")
#kable(maxb, type = "Markdown")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
maxp <- which.max(metascrubdf$profitM)
maxp <- metascrubdf[maxp,c(1,4,5)]
names(maxp) <- c("Title", "Budget", "Revenue")
ma <- rbind(maxb, maxp)
kable(ma, type = "Markdown")
```

  

### Profit vs Popularity
Correlation is .43
Popularity is not highly correlated with profit
```{r echo=FALSE, message=FALSE, warning=FALSE}
#cor(metascrubdf$profitM, metascrubdf$popularity)

plot1 <- ggplot(metascrubdf %>% filter(profitM < 1500 & popularity < 50), aes(popularity, profitM)) +
  geom_point(alpha = 0.3) + 
  geom_smooth() +
  labs(title = "Profit vs Popularity", x = "Popularity", y = "Profit") +
  theme_bw()

ggplotly(plot1)
```
Outliers: popularity: Minions, profit: Avatar
  
```{r echo=FALSE, message=FALSE, warning=FALSE}
P1 <- plot_ly(metascrubdf %>% filter(popularity >= lower & popularity < upper & popularity > 0), x=~runtime, y=~popularity, z=~profitM,color=~revenueM, mode= "markers",type = "scatter3d") %>% layout(title = "Profit vs Runtime and Popularity", scene = list(
      xaxis = list(title = "Runtime"),
      yaxis = list(title = "Popularity"),
      zaxis = list(title = "Profit")))
P1
```
  

### Profitability of genres
Action, Adventure, Comedy, and Drama top the list.
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=metascrubdf, aes(x = reorder(maingenre, -profitM), y = profitM)) +   geom_col(aes(fill=maingenre)) +
  theme(axis.text.x=element_text(angle=55, hjust=1)) +
  labs(title = "Profit by Genre", x = "Genre", y = "Profit")

```



### Genre profitability list:
```{r echo=FALSE, message=FALSE, warning=FALSE}
genprof <- aggregate(profitM ~ maingenre,metascrubdf,sum)

genprof <- genprof %>% arrange(desc(profitM))
names(genprof) <- c("Main Genre", "Profit")
kable(genprof, type = "Markdown")

```


### Profit vs Budget

Profit and profit variability increase with budget.
The top 10% budgeted movies have the most profit but also the most varibility.
```{r echo=FALSE, message=FALSE, warning=FALSE}
dta <- subset(metascrubdf,profitM<1500)
budget.break <- ntile(dta$budgetM,10)

# ggplot way
ggplot(dta, aes(as.factor(budget.break), profitM)) +
  geom_boxplot(varwidth=T, fill="plum") + 
  labs(title = "Boxplot of Profit ~ Budget", x = "Budget", y = "Profit") + 
  theme_bw()
```



### Distribution of Runtime by Genre

Mean of 111 minutes with many outliers above the upper quartile
```{r echo=FALSE, message=FALSE, warning=FALSE}

#### Interactive #####

metascrubdf1 <- metascrubdf %>% filter(maingenre > 0)
bp <- plot_ly(metascrubdf1 %>% filter(maingenre == "Action" | maingenre == "Adventure" | maingenre == "Comedy" | maingenre == "Drama" | maingenre == "Animation"), y=~runtime, color = ~as.factor(maingenre), type="box",boxpoints = "all", jitter = 0.4,pointpos = -1.8) %>% layout(title = "Runtime by Genre", xaxis = list(title = "Main Genre", showgrid = F))
bp


#mean(metascrubdf$runtime)

```



### Genre and Popularity
Popularity corresponds with profit above
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data=metascrubdf, aes(x=maingenre, y=popularity/length(popularity))) + geom_col(aes(fill=maingenre)) + theme(axis.text.x=element_text(angle=55, hjust=1)) + 
  labs(title = "Popularity by Genre", x = "Genre", y = "Popularity")
```


Here are the top ten most profitable collections:
```{r echo=FALSE, message=FALSE, warning=FALSE}
coll2 <- metascrubdf %>% group_by(Collection)%>% summarize(profit = sum(profitM))
coll2 <- as.data.frame(coll2)
coll <- coll2[with(coll2,order(-coll2$profit)),]
coll <- filter(coll, coll$profit<10000)
kable(head(coll,10), type = "Markdown")
```


```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
cast_data1 <- cast_data %>% filter(gender != "Unspecified")
pct_male <- select(cast_data1, gender, id)  %>% group_by(movie_id) %>% summarise_at(vars(gender), list(rating = count(gender == "Male")/nrow(movie_id)))

(cast_data1 %>% filter(movie_id == 862 & gender == "Male") %>% count())  / (nrow(cast_data1 %>% filter(movie_id == 862)) )

group_by(df, group) %>% mutate(percent = value/sum(value))

test <-select(cast_data1, movie_id, gender) %>% group_by(movie_id) %>%
  mutate(pct_male = ((sum(gender == "Male"))/(sum(gender == "Male" | gender == "Female"))))

pctMale <- unique(test)

movies.pctMale <- merge(x = movies.orig, y = pctMale, by.x = "id", by.y = "movie_id")


ggplot(movies.pctMale %>% filter(vote_count > 29), aes(pct_male, vote_average)) + geom_point(alpha = 0.3) + geom_smooth()

ggplot(metascrubdf %>% filter(maingenre == "Action" | maingenre == "Adventure" | maingenre == "Comedy" | maingenre == "Drama" | maingenre == "Animation"), aes(x = maingenre, y = runtime)) +
  geom_boxplot(aes(col = maingenre)) +
  labs(title = "Runtime by Genre", x = "Main Genre", y = "Runtime", subtitle = "Top 5 most profitable genres") +
  theme_bw()

ggplot(data=comps.budget, aes(x = reorder(production_companies, -budget), y = budget)) + geom_col(aes(fill=production_companies)) + 
  theme(axis.text.x=element_text(angle=55, hjust=1)) +
  labs(title = "Production Companies by Money Spent", x = "Production Companies", y = "Money Spent")



ggplot(data=comps, aes(x = reorder(production_companies, -revenue), y = revenue)) + geom_col(aes(fill=production_companies)) + 
  theme(axis.text.x=element_text(angle=55, hjust=1)) +
  labs(title = "Production Companies by Revenue", x = "Production Companies", y = "Revenue")
```
