---
title: "Munging Movies"

date: "11/15/2019"
output: html_document
---

# MICHAEL NEW CODE FROM HERE*****************



# loading libraries
```{r}
library(readr)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(lubridate)
```



# ORIGINAL METADATA

## original metadata with some dictionary form columns
```{r}
metadata <- read_csv("the-movies-dataset/movies_metadata.csv")
```

## show proportion of missing values in each column
```{r}
na.prop2 <- apply(metadata,2,function(x) sum(is.na(x)/length(x)))
signif(na.prop2,4)

```

# revenue over 100M over time

```{r}
nona <- metadata %>% filter(metadata$revenue>100000000) 
plot(nona$revenue ~ nona$release_date)
#plot(metadata$revenue ~ metadata$release_date)
```




# SCRUBBED DATA CONTAINING $ INFORMATION


# Using a scrubbed version of the metadata with nonzero $ info,converting to dataframe, and attaching
```{r}
metascrub <- read_csv("the-movies-dataset/MetaScrubbedMoney.csv")
metascrubdf <- as.data.frame(metascrub)
attach(metascrubdf)
```


# CHECKING DATA STRUCTURE

## Viewing head structure of dataframe
```{r}
head(metascrubdf)
```

## summary information
```{r}
summary(metascrubdf)
```

## show proportion of missing values in each column
```{r}
na.prop <- apply(metascrubdf,2,function(x) sum(is.na(x)/length(x)))
na.prop

```


# MONETARY ANALYSIS


## Average budget, revenue, and profit
```{r}
sapply(metascrubdf[,c(4,5,6)],mean)
```




## correlations
```{r}
cor(metascrubdf[,c(3:6,8)])
```


# REVENUE VS BUDGET


## revenue vs. budget
Correlation is .721892
```{r}
plot(revenueM ~ budgetM, data=metascrubdf)
cor(revenueM,budgetM)
```

# 2 main outliers
```{r}
maxb <- which.max((budgetM))
metascrubdf[maxb,c(1,4,5)]
```

```{r}
maxp <- which.max(profitM)
metascrubdf[maxp,c(1,4,5)]
```

Outliers:
  budget: Pirates of the Caribbean: On Stranger Tides
  revenue: Avatar


# Linear model for revenue vs. budget
revenue = -3.34 + 3.02 (budget)
Need to spend money to make money
```{r}
mdl <- (lm(revenueM ~ budgetM, data=metascrubdf))
coef(mdl)

```


# PROFIT VS POPULARITY


# profit vs popularity
Correlation is .4274831
Popularity is not highly correlated with profit
```{r}
plot(profitM ~ popularity, data=metascrubdf)
cor(profitM,popularity)
```
Outliers:
  popularity: Minions
  profit: Avatar
  
  

## Removing the top 6 outliers for profit vs popularity
Slightly lower correlation.
For the handful of highly popular movies between 100 and 250 poupularity, profit decreases with popularity.
```{r}
temp <- subset(metascrubdf, ((profitM<1500) & (popularity<250)))
plot(profitM ~ popularity, data=temp)
cor(temp$profitM,temp$popularity)

```


# GENRE PROFITABILITY


## Profitability of genres
Action, Adventure, Comedy, and Drama top the list.
```{r}
ggplot(data=metascrubdf, aes(x=maingenre, y=profitM)) + geom_col(aes(fill=maingenre))
```

## genre profitability list

```{r}
genprof <- aggregate(profitM ~ maingenre,metascrubdf,sum)
genprof
#plot(genprof$profitM,genprof$maingenre)
```


# PROFIT VS BUDGET


## boxplots for profit quantiles without top 3 outliers
Profit and profit variability increase with budget
The top 10% budgeted movies have the most profit but also the most varibility
```{r}
dta <- subset(metascrubdf,profitM<1500)
budget.break <- ntile(dta$budgetM,10)
boxplot((profitM ~ budget.break),data = dta)

```



# RUNTIME INFORMATION

## runtime information
mean of 111 minutes with many outliers above the upper quartile
```{r}
boxplot(runtime)
hist(runtime)
mean(runtime)
```



# which genres of movies are most popular
Popularity corresponds with profit above

```{r}
ggplot(data=metascrubdf, aes(x=maingenre, y=popularity/length(popularity))) + geom_col(aes(fill=maingenre))
```


## top 10 profitable collections
```{r}

coll2 <- metascrubdf %>% group_by(Collection)%>% summarize(profit = sum(profitM))
coll2 <- as.data.frame(coll2)
coll <- coll2[with(coll2,order(-coll2$profit)),] %>% filter(coll$profit<10000)
head(coll,10)
```

```{r}
scatter.smooth(coll$profit)
```









