---
title: "MovieSetsProject"
author: "Dominic Ventura"
date: "11/15/2019"
output: html_document
---


Load packages:
```{r}
setwd("/Users/dominicventura/Data Munging/Movies")

#install.packages("lubridate")
library(readr)
library(tidyverse)
library(lubridate)
library(jsonlite)
library(data.table)
```


Read data:
```{r}
movies <- read_csv("movies_metadata.csv")
credits <- read.csv("credits.csv", stringsAsFactors = FALSE)
keywords <- read_csv("keywords.csv")
links <- read_csv("links.csv")
#ratings <- read_csv("ratings.csv")
```


Check NAs:
```{r}
str(movies)
sum.NA.total <- movies %>%
  select(everything()) %>%
  summarise_all(funs(sum(is.na(.))))
PropOfNA.total <- (sum.NA.total/nrow(movies))*100
SumNA.prop <- PropOfNA.total %>% map_dbl(sum)
kable(SumNA.prop, format = "markdown")
```


# EDA
```{r}
# freq table for # movies per year:
movies_byYear <- movies %>% 
  group_by(year(release_date)) %>% 
  count()
# barplot of # movies per year:
ggplot(movies_byYear, aes(`year(release_date)`, n)) + geom_bar(stat="identity", width = 0.5)
# set movies with budget 0 to NA
movies$budget[movies$budget == 0] <- NA 
summary(movies$budget)

```



# to unpack genres column:
```{r}
metadata <- fread("movies_metadata.csv", select=c('adult', 'genres', 'release_date', 'original_language', 'original_title', 'id', 'imdb_id'), fill=T)

metadata <- metadata[!is.na(as.integer(id))]
metadata <- metadata[, genre := gsub("\'","\"", metadata$genre)]

metadata[, c('original_language', 'adult', "imdb_id") := NULL]

genres <- metadata[, unlist(lapply(genre, fromJSON), recursive=F)['name'], by=id]
sorted.genres <- genres[, .N, by=name][order(-N)]
sorted.genres[1:20]
genres[, dummy := 1]
encoded.genres <- dcast(na.omit(genres[,.SD[1:3], by=id])[name %in% sorted.genres[1:20, name]], 
                        id ~  name, value.var='dummy', fill=0)
data.w.genre <- merge(encoded.genres,
      metadata[, .(id, original_title, release_date)], 
      all.x=T, by="id")
```


```{r}
#install.packages("tm")  # for text mining
#install.packages("SnowballC") # for text stemming
#install.packages("wordcloud") # word-cloud generator 
#install.packages("RColorBrewer") # color palettes
# install.packages(c("tm", "SnowballC", "wordcloud", "RColorBrewer", "RCurl", "XML"))


library("tm")
library("SnowballC")
library("wordcloud")
library("RColorBrewer")



keywordsDS <- data.frame(keywords)

# Create the corpus of data
keywordsDS.corpus <- Corpus(VectorSource(keywordsDS$keywords))

# Cleaning data. This includes removing stopwords, numbers, whitespace, etc. and converting the corpus into a plain text document.

keywordsDS.Clean <- tm_map(keywordsDS.corpus, PlainTextDocument)
keywordsDS.Clean <- tm_map(keywordsDS.corpus, tolower)
keywordsDS.Clean <- tm_map(keywordsDS.Clean, removeNumbers)
keywordsDS.Clean <- tm_map(keywordsDS.Clean, removeWords, stopwords("english"))
keywordsDS.Clean <- tm_map(keywordsDS.Clean, removeWords, c("name"))
keywordsDS.Clean <- tm_map(keywordsDS.Clean, removeWords, c("film"))
keywordsDS.Clean <- tm_map(keywordsDS.Clean, removeWords, c("director"))
keywordsDS.Clean <- tm_map(keywordsDS.Clean, removePunctuation)
keywordsDS.Clean <- tm_map(keywordsDS.Clean, stripWhitespace)
keywordsDS.Clean <- tm_map(keywordsDS.Clean, stemDocument)

wordcloud(words = keywordsDS.Clean, min.freq = 1,
          max.words=100, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))

# Build a term-document matrix. Document matrix is a table containing the frequency of the words. Column names are words and row names are documents.
dtm <- TermDocumentMatrix(keywordsDS.Clean)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
head(d, 20)

# The frequency of the first 10 frequent words plotted
barplot(d[1:20,]$freq, 
        las = 2, 
        names.arg = d[1:20,]$word,
        col ="lightyellow", 
        main ="Most Frequent Words",
        ylab = "Word Frequencies")

# Analyze the association between frequent terms (i.e., terms which correlate)
findAssocs(dtm, terms = "love", corlimit = 0.3)
findAssocs(dtm, terms = "music", corlimit = 0.3)
findAssocs(dtm, terms = "war", corlimit = 0.3)
findAssocs(dtm, terms = "novel", corlimit = 0.3)

```

```{r}
require(stats)

pop.title <- data.frame(movies$popularity, movies$title)
ordered.by.pop <- pop.title[order(-pop.title$movies.popularity),]
head(ordered.by.pop, 10)

budge.rev <- (data.frame(movies$title, movies$revenue, movies$budget))
head(budge.rev, 10)

# find which movies made/lost most money and if difference btwn two is same as budget, we remove it.

n <- nrow(budge.rev)
names(budge.rev) <-  c("Title", "Revenue", "Budget")

diff <- data.frame(matrix(0, nrow = n, ncol=1))

# For loop to take difference of budget and revenue
                   
for (i in 1:n) {
  gross <- budge.rev[i, 2] - budge.rev[i, 3]
  diff[i,1] <- gross

}

# Binding new gross column to data frame
new.budge.rev <- cbind(budge.rev, diff)
names(new.budge.rev) <- c("Title", "Revenue", "Budget", "Gross Revenue")


# Filtering out movies with 0's, or equal to budget/revenue
better.budge.rev <- filter(new.budge.rev, new.budge.rev$matrix.0..nrow...n..ncol...1. != new.budge.rev$Revenue & new.budge.rev$matrix.0..nrow...n..ncol...1. != -new.budge.rev$Budget)


# Ordered by highest grossing
ordered.by.gross <- better.budge.rev[order(-better.budge.rev$matrix.0..nrow...n..ncol...1.),]

names(ordered.by.gross) <- c("Title", "Revenue", "Budget", "Gross")




```




```{r}
# Plot (histogram, density, box/violin) runtime and average votes to see whether liked shorter movies or longer movies (distribution?)

# Did runtime affect how watchers enjoyed movie?
library(ggplot2)
library(data.table)

ggplot(movies, aes(vote_average, runtime, color=vote_average)) + geom_point()

ggplot(movies, aes(popularity, vote_count, color=popularity)) + geom_point()

filter(movies, runtime > 1200)


```

```{r}
# Plot (histogram, density, box/violin)  votes next to gross revenue (distribution?)
```

```{r}
# Plot (histogram, density, box/violin) runtime next to gross revenue (distribution?)
``` 


## original metadata with some dictionary form columns
```{r}
metadata <- read_csv("the-movies-dataset/movies_metadata.csv")


```

## show proportion of missing values in each column
```{r}
na.prop2 <- apply(metadata,2,function(x) sum(is.na(x)/length(x)))
signif(na.prop2,4)
```

# revenue over 100M over time

```{r}
nona <- metadata %>% filter(metadata$revenue>100000000) 
plot(nona$revenue ~ nona$release_date)
#plot(metadata$revenue ~ metadata$release_date)
```




# SCRUBBED DATA CONTAINING $ INFORMATION


# Using a scrubbed version of the metadata with nonzero $ info,converting to dataframe, and attaching
```{r}
metascrub <- read_csv("/Users/dominicventura/Documents/Movies/MetaScrubbedMoney.csv")
metascrubdf <- as.data.frame(metascrub)
attach(metascrubdf)
```


# CHECKING DATA STRUCTURE

## Viewing head structure of dataframe
```{r}
head(metascrubdf)
```

## summary information
```{r}
summary(metascrubdf)
```

## show proportion of missing values in each column
```{r}
na.prop <- apply(metascrubdf,2,function(x) sum(is.na(x)/length(x)))
na.prop
```


# MONETARY ANALYSIS


## Average budget, revenue, and profit
```{r}
sapply(metascrubdf[,c(4,5,6)],mean)
```




## correlations
```{r}
cor(metascrubdf[,c(3:6,8)])
```


# REVENUE VS BUDGET


## revenue vs. budget
Correlation is .721892
```{r}
plot(revenueM ~ budgetM, data=metascrubdf)
cor(revenueM,budgetM)

```

# 2 main outliers
```{r}
maxb <- which.max((budgetM))
metascrubdf[maxb,c(1,4,5)]
```

```{r}
maxp <- which.max(profitM)
metascrubdf[maxp,c(1,4,5)]
```

Outliers:
  budget: Pirates of the Caribbean: On Stranger Tides
  revenue: Avatar


# Linear model for revenue vs. budget
revenue = -3.34 + 3.02 (budget)
Need to spend money to make money
```{r}
mdl <- (lm(revenueM ~ budgetM, data=metascrubdf))
coef(mdl)
```


# PROFIT VS POPULARITY


# profit vs popularity
Correlation is .4274831
Popularity is not highly correlated with profit
```{r}
plot(profitM ~ popularity, data=metascrubdf)
cor(profitM,popularity)
```
Outliers:
  popularity: Minions
  profit: Avatar
  
  

## Removing the top 6 outliers for profit vs popularity
Slightly lower correlation.
For the handful of highly popular movies between 100 and 250 poupularity, profit decreases with popularity.
```{r}
temp <- subset(metascrubdf, ((profitM<1500) & (popularity<250)))
plot(profitM ~ popularity, data=temp)
cor(temp$profitM,temp$popularity)
```


# GENRE PROFITABILITY


## Profitability of genres
Action, Adventure, Comedy, and Drama top the list.
```{r}
ggplot(data=metascrubdf, aes(x=maingenre, y=profitM)) + geom_col(aes(fill=maingenre))
library(ggplot2)

str(metascrubdf)

ggplot(data=metascrubdf, aes(x = maingenre, y = profitM)) + geom_col(aes(fill=maingenre)) +
  theme(axis.text.x=element_text(angle=55, hjust=1)) +
  labs(title = "Profit by Genre", x = "Genre", y = "Profit")

#metascrubdf$maingenre <- factor(metascrubdf$maingenre, levels = c("Action", "Adventure", "Animation", "Comedy", "Crime", "Documentary", "Drama", "Family", "Fantasy", "Foreign", "History", "Horror", "Music", "Mystery", "Romance", "Science Fiction", "Thriller", "TV Movie", "War", "Western", "NA"))

metascrubdf$maingenre <- as.factor(metascrubdf$maingenre)

ggplot(data=metascrubdf, aes(x = reorder(maingenre, profitM), y = profitM)) +   geom_col(aes(fill=maingenre)) +
  theme(axis.text.x=element_text(angle=55, hjust=1)) +
  labs(title = "Profit by Genre", x = "Genre", y = "Profit")
```

## genre profitability list

```{r}
genprof <- aggregate(profitM ~ maingenre,metascrubdf,sum)
genprof
#plot(genprof$profitM,genprof$maingenre)
```


# PROFIT VS BUDGET


## boxplots for profit quantiles without top 3 outliers
Profit and profit variability increase with budget
The top 10% budgeted movies have the most profit but also the most varibility
```{r}
dta <- subset(metascrubdf,profitM<1500)
budget.break <- ntile(dta$budgetM,10)
boxplot((profitM ~ budget.break),data = dta)

```



# RUNTIME INFORMATION

## runtime information
mean of 111 minutes with many outliers above the upper quartile
```{r}
boxplot(runtime)
hist(runtime)
mean(runtime)
```



# which genres of movies are most popular
Popularity corresponds with profit above

```{r}
ggplot(data=metascrubdf, aes(x=maingenre, y=popularity/length(popularity))) + geom_col(aes(fill=maingenre))
```


## top 10 profitable collections
```{r}
coll2 <- metascrubdf %>% group_by(Collection)%>% summarize(profit = sum(profitM))
coll2 <- as.data.frame(coll2)
coll <- coll2[with(coll2,order(-coll2$profit)),] %>% filter(coll$profit<10000)
head(coll,10)
```

```{r}
install.packages("plyr")
library(plyr)


production <- read_csv("/Users/dominicventura/Documents/Movies/Production.csv")
production <- data.frame(production)


comps <- aggregate(revenue ~ production_companies, data=production, FUN=sum) %>% 
  filter(revenue > 10000000000)

comps[order(-comps$revenue),]

ggplot(data=comps, aes(x = reorder(production_companies, -revenue), y = revenue)) + geom_col(aes(fill=production_companies)) + 
  theme(axis.text.x=element_text(angle=55, hjust=1)) +
  labs(title = "Production Companies by Revenue", x = "Production Companies", y = "Revenue")


comps.budget <- aggregate(budget ~ production_companies, data=production, FUN=sum) %>%
  filter(budget > 2000000000)

ggplot(data=comps.budget, aes(x = reorder(production_companies, -budget), y = budget)) + geom_col(aes(fill=production_companies)) + 
  theme(axis.text.x=element_text(angle=55, hjust=1)) +
  labs(title = "Production Companies by Money Spent", x = "Production Companies", y = "Money Spent")

```

```{r}


#num.of.movies <- sort(table(production$production_companies), decreasing = TRUE) 
#head(num.of.movies, 10)




#str_split(movies$production_companies, ",", 4, simplify = TRUE)


cast_text <- credits$cast
parse_cast <- function(cast_text) {
  text <- 
    sub("\\£\\$\\£", "",    #drop the first one
        gsub("\'|\"|\\[|\\]|\\{|\\}|,", "",
             gsub("\'cast_id\':|\'character\':|\'credit_id\':|\'gender\':|\'id\':|\'name\':|\'order\':|\'profile_path\':", 
                  "£$£", 
                  cast_text)
        )
    )
  
  data <- data.frame(matrix(
    trimws(unlist(
        strsplit(text, "\\£\\$\\£")
        )), 
    ncol = 8, byrow = TRUE), stringsAsFactors = FALSE)
  
  names(data) <- c("cast_id", "character", "credit_id", "gender", "id", "name", "order", "profile_path")
  data$order <- as.numeric(data$order)
  
  return(data %>% select(cast_id, gender, id, name, order))
}

num_batches <- 50
batch_interval <- ceiling(nrow(credits) / num_batches)
batch_brackets <- 0:num_batches * batch_interval
batch_brackets[num_batches+1] <- nrow(credits)
batch_brackets[1] <- 1

cast_data <- data.frame()
for(b_i in 1:num_batches){
  
  batch_data <- data.frame()
  for(b2_i in batch_brackets[b_i]:(batch_brackets[b_i+1]-1)){
    #print(b2_i)
    this_cast <- parse_cast(credits$cast[b2_i])
    if(nrow(this_cast) > 0){
      this_cast$movie_id <- as.character(credits$id[b2_i])
      batch_data <- rbind(batch_data, this_cast)
    }
  }
  cast_data <- rbind(cast_data, batch_data)
  print(paste0("    done batch ", b_i))
}


```
```{r}

cast_data$gender[cast_data$gender == 0] <- "Unspecified"
cast_data$gender[cast_data$gender == 1] <- "Female"
cast_data$gender[cast_data$gender == 2] <- "Male"

ggplot(cast_data, aes(factor(gender), fill=factor(gender))) + 
  geom_bar()

getwd()
write.csv(cast_data, file="castData.csv")

castDataCast <- read_csv("/Users/dominicventura/Data Munging/Movies/castData.csv")
```

```{r}
cast_text_crew <- credits$crew
parse_cast <- function(cast_text_crew) {
  text <- 
    sub("\\£\\$\\£", "",    #drop the first one
        gsub("\'|\"|\\[|\\]|\\{|\\}|,", "",
             gsub("\'credit_id\':|\'department\':|\'gender\':|\'id\':|\'job\':|\'name\':|\'profile_path\':", 
                  "£$£", 
                  cast_text_crew)
        )
    )
  
  data <- data.frame(matrix(
    trimws(unlist(
        strsplit(text, "\\£\\$\\£")
        )), 
    ncol = 7, byrow = TRUE), stringsAsFactors = FALSE)
  
  names(data) <- c("credit_id", "department", "gender", "id", "job", "name", "profile_path")
  #data$order <- as.numeric(data$order)
  
  return(data %>% select(name, gender, id, job))
}

num_batches <- 50
batch_interval <- ceiling(nrow(credits) / num_batches)
batch_brackets <- 0:num_batches * batch_interval
batch_brackets[num_batches+1] <- nrow(credits)
batch_brackets[1] <- 1

cast_data <- data.frame()
for(b_i in 1:num_batches){
  
  batch_data <- data.frame()
  for(b2_i in batch_brackets[b_i]:(batch_brackets[b_i+1]-1)){
    #print(b2_i)
    this_cast <- parse_cast(credits$crew[b2_i])
    if(nrow(this_cast) > 0){
      this_cast$movie_id <- as.character(credits$id[b2_i])
      batch_data <- rbind(batch_data, this_cast)
    }
  }
  cast_data <- rbind(cast_data, batch_data)
  print(paste0("    done batch ", b_i))
}

```



```{r}
# cast_data_crew <- read_csv("/Users/dominicventura/Data Munging/Movies/castDataCrew.csv")

cast_data_crew$gender[cast_data$gender == 0] <- "Unspecified"
cast_data_crew$gender[cast_data$gender == 1] <- "Female"
cast_data_crew$gender[cast_data$gender == 2] <- "Male"

#write.csv(cast_data, file="castDataCrew.csv")

directors <- cast_data_crew %>%
  filter(job == "Director")

ggplot(directors, aes(factor(gender), fill=factor(gender))) + geom_bar() + labs(title = "Directors by Gender", x = "Gender", y = "Count")

#unique(cast_data$job)

screenplay <- cast_data_crew %>%
  filter(job == "Screenplay")

ggplot(screenplay, aes(factor(gender), fill=factor(gender))) + geom_bar() + labs(title = "Screenplay Writers by Gender", x = "Gender", y = "Count")


producer <- cast_data_crew %>%
  filter(job == "Producer")

ggplot(producer, aes(factor(gender), fill=factor(gender))) + geom_bar() + labs(title = "Producers by Gender", x = "Gender", y = "Count")

EP <- cast_data_crew %>%
  filter(job == "Executive Producer")

ggplot(EP, aes(factor(gender), fill=factor(gender))) + geom_bar() + labs(title = "EPs by Gender", x = "Gender", y = "Count")

editor <- cast_data_crew %>%
  filter(job == "Editor")

ggplot(editor, aes(factor(gender), fill=factor(gender))) + geom_bar() + labs(title = "Editors by Gender", x = "Gender", y = "Count")


```

```{r}
order01 <- castDataCast %>%
  filter(order == 0 | order == 1)

  
topBilling <- order01 %>%
  distinct() %>%
  count(name)

names(topBilling) <- c("Name", "Count")

topBilling <- topBilling[order(-topBilling$Count),]
topBilling
  
topBilling10 <- head(topBilling[order(-topBilling$n),], 10)


ggplot(data=topBilling10, aes(x = reorder(name, -n), y = n)) + geom_col(aes(fill=name)) + 
  theme(axis.text.x=element_text(angle=55, hjust=1)) +
  labs(title = "Top Billing by Actor", x = "Actor", y = "Count")



```

```{r}
billingGender <- order01 %>%
  filter(order == 0 | order == 1) %>%
  count(gender)

billingGender

ggplot(data=billingGender, aes(x = reorder(gender, -n), y = n)) + geom_col(aes(fill=gender)) + 
  theme(axis.text.x=element_text(angle=55, hjust=1)) +
  labs(title = "Top Billing by Gender", x = "Gender", y = "Count")
```





